The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2024-06-19 analysis/paper/paper.qmd
Remove unnecessary call to data
Not needed according to R Packages 2e, Ch.7, p.102
Update paper

---
title: "The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach"
author:
  - name: Claudiu Forgaci
    correspondence: "yes"
    email: C.Forgaci@tudelft.nl
    orcid: 0000-0003-3218-5102
    affiliations: 
      - ref: tud
affiliations:
  - id: tud
    name: Delft University of Technology
    address: Julianalaan 134, 2628 BL, Delft, Zuid-Holland, The Netherlands
title-block-published: "Last updated"  
date: now
date-format: long
format:
  html: default
  # docx:
  #   reference-doc: ../templates/template.docx
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  This paper aims to describe the legacy of the European post-Master of Urbanism 
  (EMU) that ran between 2005 and 2021 at TU Delft. To that end, it reports on the
  analysis of the graduation theses produced over the duration of the program.
  The text mining approach allows for a close scrutiny of the latent thematic patterns, 
  and evolutions thereof, found in the full texts of the theses. Results reveal 5 topics
  that are to a large extent consistent with the set-up of the program in two thematic 
  semesters at TU Delft and three possible exchange semester locations, each characterised 
  by a specific approach. The paper discusses the relevance of the findings for the
  state of the profession and education in the field. Set up as a research
  compendium accompanied by a published FAIR dataset and an interactive dashboard, 
  the research is fully reproducible and enables further analysis. 
keywords: |
  urbanism; education; post-master; text mining; topic modeling
highlights: |
  These are the highlights. 
---

@@

# Load data ----
data(emu_metadata)  # Thesis metadata
data(emu_raw)       # Raw full text data
data(emu_clean)     # Cleaned full text data

@@

The research presented in this paper was prompted by the closure of the European post-Master of Urbanism (EMU) at TU Delft in 2021. The EMU is an advanced program started jointly by TU Delft, KU Leuven, UPC Barcelona and Università IUAV di Venezia in 2005. The program distinguished itself internationally as an advanced post-master open only to applicants with a prior MSc degree and practical experience in urbanism or a related field. With a combined research and design curriculum, it provided a bridge towards both advanced practice and research trajectories after graduation. All students followed an exchange semester at one of the other three program universities, allowing them to have a broad, international view on the field of urbanism. Although located in Europe and named "European", its geographic scope was global, visible especially at the stage of graduation, considering that a large majority of students were of non-European origin and chose to work on a case from their home location.

@@

## A text mining approach {#textmin}

@@

Acknowledging that opportunity, this paper leverages topic modeling, a branch of Natural Language Processing, as an unsupervised machine learning method to reveal latent semantic structure and relationships between latent topics in a large document collection [@muchene2021]. Latent Dirichlet Allocation (LDA) is broadly adopted in topic modeling [@muchene2021] as it uses a soft clustering algorithm in which objects can be part of multiple clusters (mixed membership approach) as opposed to hard clustering algorithms (single membership approaches) [@rüdiger2022]. LDA uses as input a bag-of-words (BOW) model, more generally called the vector space model (VSM), in which documents in a corpus are decomposed into two low-dimensional matrices for document-topic distribution and word-topic distribution, respectively. Although topic modeling in general automates the process of identifying topics through clustering, those topics need to be manually interpreted and labelled after the application of the algorithm [@rüdiger2022]. Moreover, arguing for the right algorithm and a suitable number of clusters remains a challenge [@rüdiger2022].

@@

# Methods {#methods}

@@

## Data collection {#datacol}

@@

theses_overview <- theses_all_per_year |> 
  select(grad_year, n_total, n_permissions) |> 
  mutate(grad_year = as.factor(grad_year),
         `% of all theses per period` = scales::label_percent()(round(n_permissions / n_total, 4))) |>
  bind_rows(data.frame(grad_year = "2007-2021", 
                       n_permissions = nrow(emu), 
                       n_total = sum(theses_all_per_year$n_total),
                       `% of all theses per period` = scales::label_percent()(nrow(emu) / sum(theses_all_per_year$n_total)),
                       check.names = FALSE)) |> 
  filter(!is.na(grad_year)) |> 
  rename(`No. theses` = n_total,
         `Graduation year` = grad_year,
         `No. theses included` = n_permissions)

@@

The dataset consists of full texts and metadata of a subset of EMU theses written in English. The full texts are available as PDF documents with unstructured layouts typical to an urbanism project in which text and various types of graphics are combined @fig-pdf-example. The full-text documents were obtained either from the TU Delft Education Repository upon request, or from the authors via email if not available on the repository. Metadata was downloaded from the TU Delft Education Repository or obtained from the EMU program coordination records if unavailable on the repository. Even though repository entries are openly accessible, licenses were not specified and thus it was unclear if the content of the theses can be mined and if the resulting data can be published. Considering the relatively small size of the corpus and the possibility of directly contacting the alumni network, permissions were requested individually. As a result, of the total of `r n_theses_all` theses presented in the EMU program, full-text documents were obtained for `r n_theses_pdf` of which permissions were granted for `r n_theses`, that is, `r perc_theses_included` of all theses. Only PDFs for which permission was obtained were included in the corpus. The final corpus contains at least one thesis for each year of graduation between `r start_year`-`r end_year` as shown in @tbl-data-selection. @fig-map-theses shows the geographic spread of the theses included in the analysis.

@@

```{r}
#| label: tbl-data-selection
#| tbl-cap: Overview of theses included in the analysis
#| tbl-cap-location: top

theses_overview |>
  kableExtra::kable(align = "lrrr")
```

```{r}
#| label: fig-map-theses
#| fig-cap: "Location of EMU theses included in the analysis"
#| warning: false

# Show thesis locations on world map
visualize_thesis_locations(emu)
```

## Pre-processing {#prepro}

@@

1.  Text was extracted from the original PDF documents with the the PyMuPDF Python package [@pymupdf/2024]. During extraction, the front matter (cover page, colophon, table of contents, acknowledgements) and back matter (references, acknowledgements, appendices) were excluded. The page numbers used to exclude the front matter and back matter were manually identified and used as input for page selection during text extraction. The extracted text was stored in the `text_raw` column of the dataset.
2.  The extracted text was cleaned in R of line breaks, hyphenation, occurrences of the thesis title, author name, and page numbers. The cleaned text data was added to each thesis in a new column `text_clean`.
3.  From the metadata only the title, graduation year, exchange semester and location of the case studied in the theses were kept. Column names were shortened and updated to follow the naming conventions used in the dataset.

@@

tibble::tibble(Column = names(emu), 
               Type = purrr::map(emu, class),
               Completeness = 
                 purrr::map(emu, is.na) |> 
                 purrr::map(mean) |> 
                 purrr::map(\(x){1 - as.numeric(x)}) |> 
                 purrr::map(scales::label_percent())) |>
  kableExtra::kable(align = "llr")

@@

## Text mining {#infoext}

<!-- TODO Add intro to this section -->

Text mining is a process of extracting information from unstructured data following the steps of tokenisation, construction of a Document-Term Matrix (DTM), fitting a topic model and visualisation of the results.

### Tokenization

@@

### Document-term matrix

@@

### Topic model

@@

| University    | Semester        | Topic                                            |
|-----------------|-----------------|--------------------------------------|
| TU Delft      | Fall semester   | Urban region networks                            |
|               | Spring semester | Constructing the sustainable city                |
| IUAV Venice   | Fall semester   | Territories of dispersion: Situations, Scenarios |
| KU Leuven     | Fall semester   |                                                  |
| UPC Barcelona | Fall semester   | Urban Transformations                            |

@@

@@

## K = 5 considering the 2 TUD semesters and the 3 exchange semesters combined 
lda <- topicmodels::LDA(dtm, k = 3, method="Gibbs",
                        control = list(seed = 2023, iter = 500, verbose = 100))

@@

# Add pseudo-names to topics based on the top 5 words in each topic
topic_names_5 <- name_topics(beta)

@@

# Results

@@

## Topics

@@

# TODO Replace this with a table containing topics, documents, and theta
#      sorted by theta in decreasing order
# What is the dominant topic in each document?
topic_prop <- as.data.frame(theta)
colnames(topic_prop) <- paste0(paste0("Topic ", 1:length(topic_names_5), ": "), topic_names_5)

@@

### Topic 1: `r topic_names_5[1]`

@@

### Topic 2: `r topic_names_5[2]`

@@

### Topic 3: `r topic_names_5[3]`

@@

### Topic 4: `r topic_names_5[4]`

@@

### Topic 5: `r topic_names_5[5]`

@@

### Integrated projects

@@

## The dynamics of thesis topics

@@

topic_prop_per_year <- theta[1:nrow(emu),]
colnames(topic_prop_per_year) <- topic_names_5

@@

# Extract documents as mixtures of topics from LDA model
documents <- tidy(lda, matrix = "gamma")

@@

# Discussion

@@

# Conclusion {#conclusion}

# Acknowledgements

@@

# References

>>>>

---
title: "The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach"
author:
  - name: Claudiu Forgaci
    correspondence: "yes"
    email: C.Forgaci@tudelft.nl
    orcid: 0000-0003-3218-5102
    affiliations: 
      - ref: tud
affiliations:
  - id: tud
    name: Delft University of Technology
    address: Julianalaan 134, 2628 BL, Delft, Zuid-Holland, The Netherlands
title-block-published: "Last updated"  
date: now
date-format: long
format:
  html: default
  # docx:
    # reference-doc: ../templates/template.docx
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  This paper aims to describe the legacy of the European post-Master in Urbanism 
  (EMU) that ran between 2005 and 2021 at TU Delft. To that end, it reports on the
  analysis of the graduation theses produced over the duration of the program.
  The text mining approach allows for a close scrutiny of the latent thematic patterns, 
  and evolutions thereof, found in the full texts of the theses. Results reveal 5 topics
  that are to a large extent consistent with the set-up of the program in two thematic 
  semesters at TU Delft and three possible exchange semester locations, each characterised 
  by a specific approach. The paper discusses the relevance of the findings for the
  state of the profession and education in the field. Set up as a research
  compendium accompanied by a published FAIR dataset and an interactive dashboard, 
  the research is fully reproducible and enables further analysis. 
keywords: |
  urbanism; education; post-master; text mining; topic modeling
highlights: |
  These are the highlights. 
---

@@

@@

The research presented in this paper was prompted by the closure of the European post-Master in Urbanism (EMU) at TU Delft in 2021. The EMU is an advanced program started jointly by TU Delft, KU Leuven, UPC Barcelona and Università IUAV di Venezia in 2005. The program distinguished itself internationally as an advanced post-master open only to applicants with a prior MSc degree and practical experience in urbanism or a related field. With a combined research and design curriculum, it provided a bridge towards both advanced practice and research trajectories after graduation. All students followed an exchange semester at one of the other three program universities, allowing them to have a broad, international view on the field of urbanism. Although located in Europe and named "European", its geographic scope was global (see @fig-map-theses), visible especially at the stage of graduation, considering that a large majority of students were of non-European origin and chose to work on a case from their home location.

@@

## A text mining approach {#tmapproach}

@@

Acknowledging that opportunity, this paper leverages topic modeling as an unsupervised machine learning method to reveal latent semantic structure and relationships between latent topics in a large document collection [@muchene2021]. Latent Dirichlet Allocation (LDA) is broadly adopted in topic modeling [@muchene2021] as it uses a soft clustering algorithm in which objects can be part of multiple clusters (mixed membership approach) as opposed to hard clustering algorithms (single membership approaches) [@rüdiger2022]. LDA uses as input a bag-of-words (BOW) model, more generally called the vector space model (VSM), in which documents in a corpus are decomposed into two low-dimensional matrices for document-topic distribution and word-topic distribution, respectively. Although topic modeling in general automates the process of identifying topics through clustering, those topics need to be manually interpreted and labelled after the application of the algorithm [@rüdiger2022]. Moreover, arguing for the right algorithm and a suitable number of clusters remains a challenge [@rüdiger2022].

@@

## Methods {#methods}

@@

### Data collection {#datacol}

@@

theses_overview <- theses_all_per_year |> 
  select(grad_year, n_total, n_permissions) |> 
  mutate(grad_year = as.factor(grad_year),
         `% of theses per period` = scales::label_percent()(round(n_permissions / n_total, 4))) |>
  bind_rows(data.frame(grad_year = "2007-2021", 
                       n_permissions = nrow(emu), 
                       n_total = sum(theses_all_per_year$n_total),
                       `% of theses per period` = scales::label_percent()(nrow(emu) / sum(theses_all_per_year$n_total)),
                       check.names = FALSE)) |> 
  filter(!is.na(grad_year)) |> 
  rename(`No. theses` = n_total,
         `Graduation year` = grad_year,
         `No. theses included` = n_permissions)

@@

The dataset consists of full texts and metadata of a subset of EMU theses written in English. The full texts are available as PDF documents with unstructured layouts typical to an urbanism project in which text and various types of graphics are combined @fig-pdf-example. The full-text documents were obtained either from the TU Delft Education Repository upon request, or from the authors via email if not available on the repository. Metadata was downloaded from the TU Delft Education Repository or obtained from the EMU program coordination records if unavailable on the repository. Even though repository entries are openly accessible, licenses were not specified and thus it was unclear if the content of the theses can be mined and if the resulting data can be published. Considering the relatively small size of the corpus and the possibility of directly contacting the alumni community, permissions were requested individually. As a result, of the total of `r n_theses_all` theses presented in the EMU program, full-text documents were obtained for `r n_theses_pdf` of which permissions were granted for `r n_theses`, that is, `r perc_theses_included` of all theses. Only PDFs for which permission was obtained were included in the corpus. The final corpus contains at least one thesis for each year of graduation between `r start_year`-`r end_year` as shown in @tbl-data-selection. @fig-map-theses shows the geographic spread of the theses included in the analysis.

@@

<!-- Table shown in html output -->

::: {.content-visible when-format="html"}
```{r}
#| label: tbl-data-selection-html
#| tbl-cap: Overview of theses included in the analysis
#| tbl-cap-location: top

theses_overview |>
  mutate(`% of theses per period` = formattable::color_tile("lightpink", "lightgreen")(readr::parse_number(theses_overview$`% of theses per period`))) |>
  rename("Count" = `No. theses included`, "Percentage" = `% of theses per period`) |> 
  kableExtra::kable(format = "html", align = "lrrr", escape = FALSE) |>
  kableExtra::column_spec(4, width = "3cm") |> 
  kableExtra::row_spec(nrow(theses_overview), bold = TRUE) |>
  kableExtra::kable_styling("hover", full_width = FALSE) |> 
  kableExtra::add_header_above(c(" " = 2, "Included" = 2))
```
:::

<!-- Table shown in non-html output -->

::: {.content-hidden when-format="html"}
```{r}
#| label: tbl-data-selection
#| tbl-cap: Overview of theses included in the analysis
#| tbl-cap-location: top

theses_overview |> 
  gt::gt() |> 
  gt::tab_options(table.width = "300")

```
:::

```{=html}
<!-- TODO Replace with interactive map when rendering HTML.
This should be an argument in the `visualise_thesis_locations()` function. -->
```
::: {.content-visible when-format="html"}
```{r}
#| label: fig-map-theses
#| fig-cap: "Location of EMU theses included in the analysis"
#| warning: false

# Show thesis locations on world map
visualize_thesis_locations(emu)
```
:::

::: {.content-hidden when-format="html"}
```{r}
#| label: fig-map-theses-html
#| fig-cap: "Location of EMU theses included in the analysis"
#| warning: false

# Show thesis locations on world map
visualize_thesis_locations(emu)
```
:::

### Pre-processing {#prepro}

@@

1.  Text was extracted from the original PDF documents with the the PyMuPDF Python package [@pymupdf/2024]. During extraction, the front matter (cover page, colophon, table of contents, acknowledgements) and back matter (references, acknowledgements, appendices) were excluded. The page numbers used to exclude the front matter and back matter were manually identified and used as input for page selection during text extraction. The extracted text was stored in the `text_raw` column of the dataset.
2.  The extracted text was cleaned in R of line breaks, hyphenation, occurrences of the thesis title, author name, and page numbers. The cleaned text data was added to each thesis in a new column `text_clean`.
3.  From the metadata only the title, graduation year, and location of the case studied in the theses were kept. Column names were shortened and updated to follow the naming conventions used in the dataset.

@@

tibble::tibble(Column = names(emu), 
               Type = purrr::map(emu, class),
               Completeness = 
                 purrr::map(emu, is.na) |> 
                 purrr::map(mean) |> 
                 purrr::map(\(x){1 - as.numeric(x)}) |> 
                 purrr::map(scales::label_percent())) |>
  kableExtra::kable(align = "llr")
```

@@

### Text mining {#textmin}

Text mining is a process of extracting implicit information from unstructured textual data [@jo2019]. following the steps of tokenisation, construction of a Document-Term Matrix (DTM), fitting a topic model and visualisation of the results.

#### Tokenization {#tokenization}

@@

#### Document-term matrix {#dtm}

@@

#### Topic model {#tomo}

@@

| University    | Semester        | Topic                                            |
|------------------|------------------|------------------------------------|
| TU Delft      | Fall semester   | Urban region networks                            |
|               | Spring semester | Constructing the sustainable city                |
| IUAV Venice   | Fall semester   | Territories of dispersion: Situations, Scenarios |
| KU Leuven     | Fall semester   |                                                  |
| UPC Barcelona | Fall semester   | Urban Transformations                            |

@@

```{r}
k.topics <- 2:9
folding <- c(rep(1:5, each = 13), rep(5, 3))

runonce <- function(k, fold) {
    testing.dtm <- which(folding == fold)
    training.dtm <- which(folding != fold)

    training.model <- topicmodels::LDA(dtm[training.dtm, ], k = k)
    test.model <- topicmodels::LDA(dtm[testing.dtm, ], model = training.model, control = list(estimate.beta = FALSE))

    topicmodels::perplexity(test.model)
}

res <- NULL

for (k in 2:9) {
    for (fold in 1:5) {
        res <- rbind(res, c(k, fold, runonce(k, fold)))
    }
}

total.perp <- tapply(res[, 3], res[, 1], sum)
round(total.perp)

plot(2:9, total.perp, type = "b", xlab = "Number of topics", ylab = "Perplexity")
```

@@

## K = 5 considering the 2 TUD semesters and the 3 exchange semesters combined 
lda <- topicmodels::LDA(dtm, k = 5, method="Gibbs",
                        control = list(seed = 2023, iter = 500, verbose = 100))

@@

# Add pseudo-names to topics based on the top 5 words in each topic
topic_pseudonames <- name_topics(beta, n_words = 5)

@@

## Results {#results}

@@

### Topics {#topics}

@@

# TODO Replace this with a table containing topics, documents, and theta
#      sorted by theta in decreasing order
# What is the dominant topic in each document?
topic_prop <- as.data.frame(theta)
colnames(topic_prop) <- paste0(paste0("Topic ", 1:length(topic_pseudonames), ": "), topic_pseudonames)

@@

#### Topic 1: `r topic_pseudonames[1]` {#topic1}

@@

#### Topic 2: `r topic_pseudonames[2]` {#topic2}

@@

#### Topic 3: `r topic_pseudonames[3]` {#topic3}

@@

#### Topic 4: `r topic_pseudonames[4]` {#topic4}

@@

#### Topic 5: `r topic_pseudonames[5]` {#topic5}

@@

#### Integrated projects {#intpro}

@@

### The dynamics of thesis topics {#topdyn}

@@

topic_prop_per_year <- theta[1:nrow(emu),]
colnames(topic_prop_per_year) <- topic_pseudonames

@@

# Extract documents as mixtures of topics from LDA model
documents <- tidytext::tidy(lda, matrix = "gamma")

@@

## Discussion {#discussion}

@@

## Conclusion {#conclusion}

## Acknowledgements

@@

## References
